### 2023-11-07 manuscript/manuscript.tex
fixed grammar, spelling, and table mistakes

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[numbers]{natbib}
\usepackage{amsfonts}
\usepackage{dingbat}

% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{The Impact of Extreme Weather Events on Property Insurance Pricing}
\author{Carol Li\\
    University of Connecticut
}

\begin{document}
\maketitle

\begin{abstract}
Weather catastrophes, like hurricanes, wildfires, hail storms, etc., have become more common and severe in recent years, which poses 
significant risks to homeowners and has significant financial consequences. The purpose of this research paper is to look into the 
impact of weather catastrophe frequency and severity on property insurance policy prices, and also, to identify strategies for easing 
the financial burden on homeowners. This research combines statistical analysis and data science techniques, drawing on data from 
reliable sources such as the Insurance Information Institute and Aon's 2023 Weather, Climate, and Catastrophe Insight Report. The 
research aims to provide helpful insights into the relationships between the frequency and severity of weather catastrophes and 
property insurance pricing in order to inform insurance industry practices and public policy decisions to improve resilience to 
catastrophes.
\end{abstract}

\section{Introduction}
\label{sec:intro}
In recent years, weather catastrophes have become more frequent and severe, which poses a potential threat to homeowners' safety as 
well as financial impacts. As the frequency and severity of disasters increase, the significance of the ability of homeowners to 
protect their homes using property insurance also increases. This research will focus on exploring the relationship between the 
frequency and severity of weather catastrophes and the pricing of property insurance.

Climate disasters have become more frequent and severe in recent years. These incidents not only put homeowners' safety at risk but 
also have significant economic consequences. As weather-related risks increase in frequency and severity, the ability for homeowners to 
protect their valuable assets through property insurance has also become increasingly important This paper aims to explore the 
relationship between the frequency and severity of weather hazards and property insurance policies and prices.

The importance of this study is the fact that weather disasters can carry profound amount of consequences, which affects the individual 
homeowners and the insurance industry as a whole. Homeowners are the ones who typically bear the financial burden of disasters, and face 
the finance of all the insurance premiums, deductibles, and coverage limitations at the same time. On the other hand, the insurance 
company is tasked to manage the increased risks associated with these events, which can also have a significant impact on their financial 
stability.

While previous research has examined different aspects of this complex issue, including the impact of weather risks on insurance 
prices, there are much more waiting to be discovered. Our research is based on valuable insights into the value of previous work 
\cite{hurricaneco} aimed at assets that have provided insurance pricing of weather frequency and severity. It provides a general 
understanding of how frequency and severity affect the pricing of property insurance. We are also trying to identify effective 
strategies that can reduce the financial burden of homeowners in the face of these major disasters.

This research has used several methods, including statistical analysis and data science methods. To achieve our goals, we will draw 
data from reliable sources, such as the Insurance Information Agency, National Insurance Centers for Environmental Informations, and 
Aon's 2023 Weather, Climate and Disaster Investigation Report. This study is not only timely but critically important at a time of 
increasing weather-related risks, with the goal of contributing to homeowner well-being and the stability of the property insurance 
market.

The rest of the paper is organized as follows.
The data will be presented in Section~\ref{sec:data}.
The methods are described in Section~\ref{sec:meth}.
The results are reported in Section~\ref{sec:resu}.
A discussion concludes in Section~\ref{sec:disc}.

\section{Data}
\label{sec:data}
To comprehend the impact of weather-related catastrophic events on property insurance pricing, we will begin by examining a dataset of 
billion-dollar disasters. This dataset provides crucial information on event frequency, financial cost, and other key parameters. Our 
data sources include the National Centers for Environmental Information (NCEI)\cite{ncei}, the Insurance Information Institute 
(III)\cite{iii}, Aon\cite{aon} and the National Association of Insurance Commissioners (NAIC)\cite{naic}, all are recognized for their 
reliability in documenting catastrophic events and their associated economic and insured losses.

Our primary dataset, obtained from the National Centers for Environmental Information (NCEI)\cite{ncei}, spans between the years from 
1980 to 2023. We have organized the data into various temporal segments to facilitate analysis. Table \ref{tab:bil_dol_disasters} 
summarizes key statistics from this dataset, which provides insights into the frequency of catastrophic events, total financial cost, and 
associated fatalities. These temporal segments, ranging from the 1980s to the present day, enable us to assess trends over time and 
help to identify potential patterns.

\begin{table}[h]
    \caption{NEIC: Billion-Dollar Disasters Data}
    \label{tab:bil_dol_disasters} 
    \centering
    \resizebox{\columnwidth}{!}{\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
        \hline
        Time Period & Billion-Dollar Disasters & Events/Year & Cost & Percent of Total Cost & Cost/Year & Deaths & Deaths/Year \\
        \hline
        1980s (1980-1989) & 33 & 3.3 & \$213.6B & 8.1\% & \$21.4B & 2,994 & 299 \\
        1990s (1990-1999) & 57 & 5.7 & \$326.8B & 12.4\% & \$32.7B & 3,075 & 308 \\
        2000s (2000-2009) & 67 & 6.7 & \$604.2B & 22.9\% & \$60.4B & 3,102 & 310 \\
        2010s (2010-2019) & 131 & 13.1 & \$967.4B & 36.7\% & \$96.7B & 5,227 & 523 \\
        Last 5 Years (2018-2022) & 90 & 18.0 & \$623.0B & 23.6\% & \$124.6B & 1,751 & 350 \\
        Last 3 Years (2020-2022) & 60 & 20.0 & \$456.0B & 17.3\% & \$152.0B & 1,460 & 487 \\
        Last Year (2022) & 18 & 18.0 & \$178.8B & 6.8\% & \$178.8B & 474 & 474 \\
        All Years (1980-2023)* & 372 & 8.5 & \$2,635.1B & 100.0\% & \$59.9B & 16,231 & 369 \\
        \hline
    \end{tabular}}
    \cite{ncei}
\end{table}

To further examine the financial implications of natural disasters, we turn to the Insurance Information Institute (III)\cite{iii}. 
Table \ref{tab:natural_cat_losses} for the year 2022 provides a breakdown of catastrophic events by peril, including data on the 
number of events, fatalities, economic losses, and insured losses. This granular information is essential for understanding the 
varying impacts of different types of natural catastrophes and their associated costs.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{NCEI US disaster freq.pdf}
    \label{fig:disaster_freq}
    \cite{ncei}
\end{figure}

This chart \ref{fig:disaster_freq} shows the percentage frequency of years from 1980-2023 that experienced different numbers of billion-dollar disasters. It 
looks at years with 1 or more, up to 5 or more billion-dollar disaster events. The chart helps to characterize the concentration and 
frequency of multiple disasters in a single year.

\begin{table}[h]
    \caption{Natural Catastrophe Losses in the United States by Peril, 2022 (in \$ millions)}
    \label{tab:natural_cat_losses}
    \centering
    \resizebox{\columnwidth}{!}{\begin{tabular}{|l|c|c|c|c|}
        \hline
        Peril & Number of Events & Fatalities & Economic Losses (2) & Insured Losses (3) \\
        \hline
        Tropical Cyclone & 3 & 157 & \$ 96,097 & \$ 53,203 \\
        Severe Convective Storm & 62 & 49 & \$ 37,232 & \$ 29,306 \\
        Wildfire, Drought, Heatwave & 26 & 65 & \$ 18,093 & \$ 8,902 \\
        Winter Storm & 13 & 123 & \$ 6,223 & \$ 4,128 \\
        Flooding & 15 & 72 & \$ 7,234 & \$ 3,346 \\
        Total & 119 & 466 & \$ 164,879 & \$ 98,885 \\
        \hline
    \end{tabular}}
    \cite{iii}
\end{table}

The Insurance Information Institute (III) \cite{iii} also supplies data on insured property losses in the United States for the years 
2013-2022. Table \ref{tab:insured_prop_losses} presents both the nominal loss values when the events occurred and their equivalent 
values in 2022 dollars. Analyzing this information will allow us to assess how insured losses have evolved over the past decade.

\begin{table}[h]
    \caption{Estimated Insured Property Losses, U.S. Natural Catastrophes, 2013-2022 (in \$ billions)}
    \label{tab:insured_prop_losses}
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Year & In dollars when occurred & In 2022 dollars (2) \\
        \hline
        2013 & \$ 24.1 & \$ 31.0 \\
        2014 & \$ 23.2 & \$ 29.2 \\
        2015 & \$ 22.9 & \$ 28.8 \\
        2016 & \$ 31.6 & \$ 39.3 \\
        2017 & \$ 130.9 & \$ 158.7 \\
        2018 & \$ 60.4 & \$ 71.6 \\
        2019 & \$ 38.8 & \$ 45.2 \\
        2020 & \$ 81.0 & \$ 93.3 \\
        2021 & \$ 93.3 & \$ 102.7 \\
        2022 & \$ 98.8 & \$ 99.9 \\
        \hline
    \end{tabular}
    \cite{iii}
\end{table}

\begin{figure}[ht]
    \centering
    \label{fig:disaster_threats}
    \includegraphics[width=0.8\linewidth]{NAIC HO threats.pdf}
    \cite{naic}
\end{figure}

This figure \ref{fig:disaster_threats} shows the cumulative billion-dollar disaster costs in the United States from 1980-2023 on a year-to-date basis. It 
illustrates how annual disaster costs accumulate through the year, with key major disaster event years highlighted. The table 
provides context on total disaster costs over time.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{NAIC Property Threat by Regions.pdf}
    \label{fig:regional_disasters}
    \cite{naic}
\end{figure}

This figure \ref{fig:regional_disasters} shows the number of billion-dollar disaster events by type in the United States from 1980-2023. It breaks down event counts 
for droughts, flooding, freezes, severe storms, tropical cyclones, wildfires, and winter storms. The total disaster cost is also shown 
accumulated across years. The chart gives overview insight into disaster frequency and cost by peril.

\subsection{Key Equations}

To better understand the relationships between weather catastrophes, property insurance, and financial impact, we will first introduce 
several key equations that will act as a basis to guide our analysis:

\begin{equation}
    \label{eq:freq}
    \text{Frequency} = \frac{\text{Number of Loss Events}}{\text{Exposure Units}}
\end{equation}

The frequency equation \ref{eq:freq} calculates the frequency of loss events by dividing the number of loss events by exposure units.

\begin{equation}
    \label{eq:sev}
    \text{Severity} = \frac{\text{Total Loss Amount}}{\text{Number of Loss Events}}
\end{equation}

The severity equation \ref{eq:sev} is determined by dividing the total loss amount by the number of loss events.

\begin{equation}
    \label{eq:losscost}
    \text{Loss Cost} = \text{Frequency} \times \text{Exposure Units}
\end{equation}

The loss count equation \ref{eq:losscost} is the product of frequency and exposure units, providing insights into the overall loss count due to catastrophic events.

\begin{equation}
    \label{eq:premium}
    \text{Premium} = \lambda \cdot \text{Severity} \cdot \text{Exposure Units}
\end{equation}

The premium equation \ref{eq:premium} calculates the estimated premium based on the frequency, severity, and exposure units. This 
reflects the financial implications for homeowners and insurers.

We hope to gain a thorough understanding of the complex relationship between weather-related disasters and property insurance pricing 
by understanding these fundamental equations. These equations will serve as a basis for our data analysis, allowing us to effectively 
assess the impact of catastrophic events on property insurance while also assisting us in meeting our research objectives.

This comprehensive dataset, supported by granular data from the Insurance Information Institute \cite{iii}, serves as the foundation 
for our research. The various chronological segments, as well as the key equations, provide us with the resources that we need to 
delve deeper into the impact of weather-related catastrophes on property insurance, allowing us to effectively address our research 
objectives.

\section{Methods}
\label{sec:meth}
We will estimate the influence of natural disasters on homeowners insurance premiums using panel regression models with state and year 
fixed effects:

\begin{equation} 
    \mathrm{Premium}_{ist} = \beta_0 + \beta_1 \cdot \mathrm{Disasters}_{ist} + \gamma_i + \delta_t + \epsilon_{ist}
\end{equation}

The dependent variable is the logged average premium in state $i$ and year $t$. The main independent variables are disaster measures 
for state-year:

\begin{itemize} 
    \item Number of events 
    \item Total cost 
    \item Insured losses (also known as loss cost\ref{eq:losscost})
    \item Indicators for peril types: hurricane, flood, severe storm, winter weather, wildfire, earthquake 
\end{itemize}

Given the skewed distribution of disaster cost variables, logs are used. Fixed effects by state and year account for time-dependent 
differences between states and national trends. We begin by looking at the impact of aggregated disaster activity. The models are then 
estimated by interacting the disaster measures with peril indicators in order to compare the sensitivity across event types. Quantile 
regression was also used to determine whether or not effects differ across the premium distribution.

\section{Results}
\label{sec:resu}
\subsection{Summary Statistics}
Table 4 \ref{tab:summary} presents summary statistics for the premium and disaster data. The disaster measures show substantial variability, 
highlighting the irregular nature of extremes. Hurricane and flooding perils account for the largest share of overall cost and 
insured losses.

\begin{table}[h]
    \label{tab:summary}
    \centering
    \begin{tabular}{|l|c|c|c|}
        \hline
        & Mean & SD & Range \\
        \hline
        Premium & \$\num{959.2} & \$\num{238.5208} & \$\num{536} - \$\num{1311} \\
        Number of Disasters & MeanDisaster & SDDisaster & MinDisaster - MaxDisaster \\
        Total Cost (\$) & MeanCost & SDCost & MinCost - MaxCost \\
        Insured Losses (\$) & MeanLosses & SDLosses & MinLosses - MaxLosses \\
        \hline
    \end{tabular}
    \caption{Summary Statistics}
    \cite{statista, ncai, fema}
\end{table}

\subsection{Regression Results}
Table 5 \ref{tab:reg_results} presents results from the panel regressions. The number of disasters and total cost are significantly 
associated with higher premiums based on the log-log specification. A 10\% increase in disasters corresponds to a 14.7\% rise in 
premiums. Meanwhile, a 10\% rise in total damage leads to a 23.1\% premium increase.

\begin{table}[h]
    \caption{Regression Results}
    \label{tab:reg_results}
    \centering
    \begin{tabular}{|l|c|c|c|}
        \hline
        & (1) & (2) & (3) \\
        \hline
        Log(Premium) & 0.147$^{\ast}$ & & \\
        & (0.082) & & \\
        Log(Cost of Catastrophe Damages) & & 0.231$^{\ast}$ & 0.230$^{\ast}$ \\
        & & (0.115) & (0.115) \\
        Log(Insured Catastrophe Loss) & & & 0.147$^{\ast}$ \\    
        & & & (0.082) \\
        \hline
        State FE & \checkmark & \checkmark & \checkmark \\
        Year FE & \checkmark & \checkmark & \checkmark \\
        Observations & 19 & 19 & 19 \\
        $R^2$ & 0.160 & 0.193 & 0.160 \\
        \hline
    \end{tabular}

    \cite{statista, ncai}
  \end{table}

  The results presented in Table 6 \ref{tab:reg_peril} indicate that hurricane and flood disasters have the most 
  substantial impacts on insurance premiums. Specifically, a 10\% increase in hurricane or flood Damages are associated with 
  approximately a 0.01\% increase in premiums. Severe storms and winter weather also exhibit positive effects on premiums. However, 
  wildfires and earthquakes appear to have less influence, possibly due to their more limited geographic impact.

\begin{table}[h]
    \centering
    \caption{Disaster Effects by Peril}
    \label{tab:reg_peril}
    \begin{tabular}{|l|c|c|c|}
      \hline
      & (1) & (2) & (3) \\
      \hline
      Log(Severe Storm Cost) & 0.1776727 & & \\
      & & (0.05052779) & \\
      Log(Flood Cost) & & & \\
      & & & (0.1410256) \\
      Log(Wildfire Cost) & & & \\
      & & & \\
      \hline
      State FE & \checkmark & \checkmark & \checkmark \\
      Year FE & \checkmark & \checkmark & \checkmark \\
      Observations & 756 & 756 & 756 \\
      $R^2$ & 0.935 & 0.935 & 0.935 \\
      \hline
    \end{tabular}

    \cite{statista, ncai}
  \end{table}

\section{Discussion}
\label{sec:disc}

The study found a statistically significant link between extreme weather events and homeowner insurance pricing. According to the 
results, more frequent and severe events are associated with higher premiums, especially for hurricanes and flooding perils. This is 
consistent with previous findings on the impact of catastrophic losses on insurance markets \cite{aon}.

However, there are limitations to the data collected. Analytics may benefit from having more granular, address-level premium data to 
better pinpoint local hazards. In addition, the model fails to take account of changes in exposures and vulnerabilities over time as risk 
factors. Further studies should be conducted to mitigate the influence of climate risk data change from detailed measures, which 
increases risks.

Still, these results demonstrate an interrelation between catastrophic incidents and rising property insurance premiums - two trends 
with many causes and effects that overlap significantly. As climate change intensifies weather extremes, homeowners could face 
financial strain - potentially undermining the stability of insurance markets and leading to economic instability for policy 
interventions like subsidized insurance policies\cite{iii}, resilience incentives, or residential investments\cite{kousky} that 
might provide relief. Climate risk modeling and risk-based pricing will become even more crucial to insurers as a means of providing 
long-term financial protection from rising extremes.

\bibliography{refs}
\bibliographystyle{chicago}

\end{document}